Automating Verification of MPC Security

### 2023-06-20 popl24/pmf.tex
typos pmf.tex

We consider random tapes to be selected from a uniform distribution.
In our threat model we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ and $\flips(\prog) = F$ we will consider all $\store
\in \mems(S \cup V \cup F)$ such that
$
\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
  In some cases we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progd(\prog, \be)$ to denote the program
  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

@@

\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
to leak information probabilistically through obvservable dependencies
in views. In the basic distribution for Example \ref{example-lambda-obliv}(a)
illustrated in Figure \ref{fig-basic-distributions}, we can see that
the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
three different runs of the protocol, and in 2 out of 3 the
honest secret $\sx{1}{0}$ is 1. More precisely:
$$
(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
$$
An important point illustrated by this example is that the
probability of values in views, and not just the values themselves,
can release information to the adversary.

\paragraph{Example \ref{example-he} revisited} This standard example
of an MPC secure protocol illustrates how a functionality- in this
case addition (in the binary field)- can be implemented securely. As
mentioned previously, it is \emph{not} the case that \emph{no}
information about input secrets can be revealed to the adversary in
the MPC threat model, since the result of the functionality can itself
reveal information. For example, if the result of addition is $0$,
the adversary knows that two of the clients' secrets must be 0
and the third's must be 1, though not whose is whose- so we can
observe:
\begin{mathpar}
  (\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \}|\{ \vx{0}{output} \mapsto 0 \}) = .33
\end{mathpar}
Further, since the standard MPC threat model allows partipating
clients to be corrupted we can formally consider the impacts on
information released to the adversary by conditioning the program
distribution on corrupt secrets.  So if we assume client 2
with input secret 1 is corrupted, and we assume the output is 0, then
adversary knows with certainty that clients 1 and 3 both have input
secrets 0. We can express this as:
$$
\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{3}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 1, \vx{0}{output} \mapsto 0 \})
= 1
$$ A critical point to note is that this information revealed to the
adversary does not violate MPC security- under the assumption of
public release of the functionality result, and under the assumption
of the possible corruption of clients, this is the cost of doing
business. Rather, the concern is that the protocol does not release
any additional information through information communicated to views
of corrupted parties. Additive secret sharing as realized in
$\prog_{\ref{example-he}}$ is MPC secure in this sense- so for
example, for any $\beta_1$ and $\beta_2$ we have:
$$
\begin{array}{c}
\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
= \\
\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \vx{2}{s1} \mapsto \beta_1, \vx{2}{s3} \mapsto \beta_2, \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
= \\
.5
\end{array}
$$

>>>>

We consider random tapes to be selected from a uniform distribution.
In our threat model, we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ and $\flips(\prog) = F$ we will consider all $\store
\in \mems(S \cup V \cup F)$ such that
$
\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this, we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
  In some cases, we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progd(\prog, \be)$ to denote the program
  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

@@

\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
to leak information probabilistically through observable dependencies
in views. In the basic distribution for Example \ref{example-lambda-obliv}(a)
illustrated in Figure \ref{fig-basic-distributions}, we can see that
the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
three different runs of the protocol, and in 2 out of 3 the
honest secret $\sx{1}{0}$ is 1. More precisely:
$$
(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
$$
An important point illustrated by this example is that the
probability of values in views, and not just the values themselves,
can release information to the adversary.

\paragraph{Example \ref{example-he} revisited} This standard example
of an MPC secure protocol illustrates how a functionality- in this
case addition (in the binary field)- can be implemented securely. As
mentioned previously, it is \emph{not} the case that \emph{no}
information about input secrets can be revealed to the adversary in
the MPC threat model, since the result of the functionality can itself
reveal information. For example, if the result of addition is $0$,
the adversary knows that two of the clients' secrets must be 0
and the third's must be 1, though not whose is whose- so we can
observe:
\begin{mathpar}
  (\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \}|\{ \vx{0}{output} \mapsto 0 \}) = .33
\end{mathpar}
Further, since the standard MPC threat model allows participating
clients to be corrupted we can formally consider the impacts on
information released to the adversary by conditioning the program
distribution on corrupt secrets.  So if we assume client 2
with input secret 1 is corrupted, and we assume the output is 0, then
the adversary knows with certainty that clients 1 and 3 both have input
secrets 0. We can express this as:
$$
\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{3}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 1, \vx{0}{output} \mapsto 0 \})
= 1
$$ A critical point to note is that this information revealed to the
adversary does not violate MPC security- under the assumption of
public release of the functionality result, and under the assumption
of the possible corruption of clients, this is the cost of doing
business. Rather, the concern is that the protocol does not release
any additional information through information communicated to views
of corrupted parties. Additive secret sharing as realized in
$\prog_{\ref{example-he}}$ is MPC secure in this sense- so for
example, for any $\beta_1$ and $\beta_2$ we have:
$$
\begin{array}{c}
\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
= \\
\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \vx{2}{s1} \mapsto \beta_1, \vx{2}{s3} \mapsto \beta_2, \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
= \\
.5
\end{array}
$$
